---
title: "Palmer Penguins"
author: "Alana Unfried"
format: 
  dashboard:
    orientation: columns
    embed-resources: true
---

```{r}
#| label: setup
#| include: false

#add any packages that you need here
library(tidyverse)
library(plotly)
library(palmerpenguins)
library(ggiraph)
library(patchwork)

data(penguins)

p_clean <- penguins |>
  drop_na()

```

# Palmer Pengins

This data comes from the [Palmer Penguins R package](https://allisonhorst.github.io/palmerpenguins/)

Research Questions

1.  What is the relationship between flipper length and bill length for different species of penguins?

2.  How does bill length vary among different species of penguins?

3.  What is the distribution of body mass for each species of penguin?

# Plotly Examples

## Column {width="50%"}

This scatterplot shows the most basic form of an interactive visualization using plotly.

### Scatterplot

```{r}
#| title: Penguin Mass vs Flipper Length by Species

#scatterplot of bill length vs flipper length colored by species
#this code comes directly from the PalmerPenguins R package documentation https://allisonhorst.github.io/palmerpenguins/articles/examples.html
mass_flipper  <- ggplot(data = penguins, 
                       aes(x = flipper_length_mm,
                           y = body_mass_g)) +
  geom_point(aes(color = species, 
                 shape = species),
             size = 3,
             alpha = 0.8) +
  scale_color_manual(values = c("darkorange","purple","cyan4")) +
  labs(title = "Penguin size, Palmer Station LTER",
       subtitle = "Flipper length and body mass for Adelie, Chinstrap and Gentoo Penguins",
       x = "Flipper length (mm)",
       y = "Body mass (g)",
       color = "Penguin species",
       shape = "Penguin species") +
  theme(legend.position = c(0.2, 0.7),
        plot.title.position = "plot",
        plot.caption = element_text(hjust = 0, face= "italic"),
        plot.caption.position = "plot")

#add interactivity
ggplotly(mass_flipper )
```

## Column {width="50%"}

### Chart 2

Could also add some text here...

```{r}
#| title: Flipper Length by Species

#this code comes directly from the PalmerPenguins R package documentation https://allisonhorst.github.io/palmerpenguins/articles/examples.html

flipper_box <- ggplot(data = penguins, aes(x = species, y = flipper_length_mm)) +
  geom_boxplot(aes(color = species), width = 0.3, show.legend = FALSE) +
  geom_jitter(aes(color = species), alpha = 0.5, show.legend = FALSE, position = position_jitter(width = 0.2, seed = 0)) +
  scale_color_manual(values = c("darkorange","purple","cyan4")) +
  labs(x = "Species",
       y = "Flipper length (mm)")

#add interactivity
ggplotly(flipper_box)
```

### Text beyween charts

This chart shows a more sophisticated tool tip with custom labels.

### Chart 3


```{r}
#| title: Flipper Length by Species

#create custom labels for each point
#the <br> says to put the next text on a new line
penguins <- penguins |>
  mutate(mylabel = paste("Species:", species,
                         "<br>Flipper length (mm):", flipper_length_mm,
                         "<br>Bill length (mm):", bill_length_mm,
                         "<br>Island:", island))

#this code comes directly from the PalmerPenguins R package documentation https://allisonhorst.github.io/palmerpenguins/articles/examples.html

#to get custom labels to work, we need to map text as an aesthetic
flipper_bill <- ggplot(data = penguins,
                         aes(x = flipper_length_mm,
                             y = bill_length_mm,
                             text = mylabel)) +
  geom_point(aes(color = species, 
                 shape = species),
             size = 3,
             alpha = 0.8) +
  scale_color_manual(values = c("darkorange","purple","cyan4")) +
  labs(title = "Flipper and bill length",
       subtitle = "Dimensions for Adelie, Chinstrap and Gentoo Penguins at Palmer Station LTER",
       x = "Flipper length (mm)",
       y = "Bill length (mm)",
       color = "Penguin species",
       shape = "Penguin species") +
  theme(legend.position = c(0.85, 0.15),
        plot.title.position = "plot",
        plot.caption = element_text(hjust = 0, face= "italic"),
        plot.caption.position = "plot")


#add interactivity with custom tool tip. Cleans up variable names, stops species from repeating twice, and adds in Island information
ggplotly(flipper_bill, tooltip = c("mylabel"))
```

# ggiraph examples {orientation="rows"}

Notice the "row" orientation of this dashboard!

## Row

### 

```{r}
#| title: Highlight Species in Both Scatterplots

#This code comes from https://allisonhorst.github.io/ggiraph-demo/demo.html

#first graph
p1 <- ggplot(data = penguins, aes(x = flipper_length_mm, 
                                  y = body_mass_g)) +
  geom_point_interactive(aes(color = species, 
                             tooltip = species, 
                             data_id = species)) +
  scale_color_manual(values = c("cyan4", "darkorange", "darkorchid4")) +
  theme_minimal()

#second graph
p2 <- ggplot(data = penguins, aes(x = bill_depth_mm, 
                                  y = bill_length_mm)) +
  geom_point_interactive(aes(color = species, 
                             tooltip = species, 
                             data_id = species)) +
  scale_color_manual(values = c("cyan4", "darkorange", "darkorchid4")) +
  theme_minimal()


#combine graphs with patchwork package (p1 + p2) and make interactive with girafe function
girafe(ggobj = (p1 + p2), 
#       width_svg = 8, 
#       height_svg = 4,
       options = list(
         opts_hover(css = "fill: black;"),
         opts_hover_inv(css = "opacity: 0.1;"),
         opts_sizing(rescale = FALSE)
       ))
```

