---
title: "Select All That Apply and ChatGPT"
format: html
---

## Goal

My goal is to graph a select-all-that-apply survey question as 1 bar per response option, showing the proportion of respondents who chose each option. Here is what some example data looks like for this structure, assumign there was a select-all-that-apply question with 3 response options and 5 respondents:

```{r}
df <- tibble(
  respondent_id = 1:5,
  option_a = c(1, 0, 1, 1, 0),
  option_b = c(0, 1, 1, 0, 1),
  option_c = c(1, 1, 0, 1, 1)
)

df
```


## From Scratch

If you wanted to create this graph without using generative AI to assist you, how would you get started? Write a brief paragraph discussing how you would tackle the issue of writing this code. What would be challenging? What seems straightforward, if anything?

> Your Answer

## Code from ChatGPT

Within my ChatGPT Custom R project using model 4o, I gave the following prompt:

> how to graph a select all that apply question. each option is listed in its own binary column

Here is the response that ChatGPT returned, followed by the code:

*To graph a "select all that apply" question where each option is stored in its own binary column (e.g., 1 for selected, 0 or FALSE for not), you’ll want to:*

-   *Pivot the data longer so each response is one row per option.*

-   *Summarize counts or percentages of how often each option was selected.*

-   *Plot the summarized data using ggplot2.*

```{r}
#| label: chat gpt code

library(tidyverse)

# Example data
df <- tibble(
  respondent_id = 1:5,
  option_a = c(1, 0, 1, 1, 0),
  option_b = c(0, 1, 1, 0, 1),
  option_c = c(1, 1, 0, 1, 1)
)

# Pivot longer
df_long <- df |>
  pivot_longer(
    cols = starts_with("option_"),
    names_to = "option",
    values_to = "selected"
  ) |>
  filter(selected == 1)

# Summarize
summary_df <- df_long |>
  count(option) |>
  mutate(percent = n / sum(n) * 100)

# Plot
summary_df |>
  ggplot(aes(x = fct_reorder(option, n), y = n)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(
    title = "Selections per Option",
    x = "Option",
    y = "Number of Selections"
  ) +
  theme_minimal()
```

## Your Turn

1. Run the code above. Does it work?

> Answer here

2. How does the R code look? Does it use the standard building blocks we have learned in class?

> Answer here

3. Does the graph appear to solve the problem that I prompted? 

> Answer here

4. Open the data, or print it in the code chunk below. Write a brief explanation of what each row and column of the data represents.

> Answer here

```{r}

df

```

5. By hand/calculator, calculate the number of respondents that chose option a, option b, and option c. Do these counts match what is shown in the graph? 

> Answer here

6. Where did ChatGPT go wrong? Look at each section of the code and explain where you think ChatGPT made a mistake.

> Answer here

7. Now it's your turn. Using the same example dataset that ChatGPT generated, can you edit or write new code to correctly graph the **number** of respondents that selected each options?

```{r}

```


8. Can you edit the code to now graph the percentage of respondents that selected each response option?

```{r}

```


9. What did you learn about using ChatGPT to generate R code? How was it helpful? How did it hinder you?

> Answer Here

10. Render your .qmd file, download the resulting .html file, and submit your file to Canvas for credit. Each students should submit their own file. 
